---
title: "RMarkdown"
date: "3 4 2020"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Content

- What is R Markdown?
- Summary of the advantages
- Approach
- Syntax for the main functions
- Specific examples
- Overview
- Exercices and closure
- Refernces

## What is R Markdown?

**Markdown** = Language for text fomrmatting  

* Markdown  
* an easy to write plain text format for creatin dynamic documents and reports  
* contains "normal" text and code chunks (e.g. R, but Python, SQL, and more also possible)  
* RMD -> MD -> html, Docx or PDF  
* Formatting documents outside the analysis  

**Creating a document that can contain both text and code:**  

* When creating ("knitting"), the code is executed and displayed together with the description.  
* Only a document with out warnings/errors can be knitted.  
* Possible to export a Word, HTML or PDF file.  

## Approach
1. Open RStudio -> File -> New File -> R Markdown
2. Enter a title for the document (optionally enter an author)
3. Choose an output format
4. YAML Headline: Final choice of your output format
5. Write the markdown
6. Embed Code
7. Rendering Output  
    + RStudio: "Knit" (Ctrl+Shift+K)
    + Command line: rmarkdown:: render("input.Rmd")

## Installation
Before you start, make sure you have installed the rmarkdown package  

`install.package("rmarkdown")`

## Choose the output format
* You can shoose between HTML, PDF and Docx(Word)  
   + **HTML** is recommeneded, since everybody can read it
* You can always change your output format, in the  YAML Header

## YAML Header
### Example
title: "Exampled_rmkardown"  
author: "rmarkdown_group"  
date "3 4 2020"  
output: html_document  
  
### Output types

Type | Format | Option in YAML Header
-----|--------|------------------------
Website | HTML | output: html_document  
Document | PDF | output: pdf_document
Document (Word) | RTF | output: word_document
Presentation (beamer) | PDF | output: beamer_presentation
Presentation (ioslides) | HTML | output: ioslides_presentation



## Embed your code
**inline code **   
```{r echo=FALSE}
bq="`"
```
The mean speed of cars is (`` `r sprintf("%sr mean(cars$speed)%s", bq, bq)` ``) `r mean(cars$speed)`. 

**code chunk**   
{r cars, echo=TRUE}

```{r cars, echo=TRUE}
summary(cars)
```

