---
title: "R Markdown"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: beamer_presentation
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Content

- What is R Markdown?
- Summary of the advantages
- Approach
- Syntax for the main functions
- Specific examples
- Overview
- Exercices and closure
- Refernces

## What is R Markdown?

**Markdown** = Language for text fomrmatting  

* Markdown  
* an easy to write plain text format for creatin dynamic documents and reports  
* contains "normal" text and code chunks (e.g. R, but Python, SQL, and more also possible)  
* RMD -> MD -> html, Docx or PDF  
* Formatting documents outside the analysis  

**Creating a document that can contain both text and code:**  

* When creating ("knitting"), the code is executed and displayed together with the description.  
* Only a document with out warnings/errors can be knitted.  
* Possible to export a Word, HTML or PDF file.  

## Approach
1. Open RStudio -> File -> New File -> R Markdown
2. Enter a title for the document (optionally enter an author)
3. Choose an output format
4. YAML Headline: Final choice of your output format
5. Write the markdown
6. Embed Code
7. Rendering Output  
    + RStudio: "Knit" (Ctrl+Shift+K)
    + Command line: rmarkdown:: render("input.Rmd")

## Installation
Before you start, make sure you have installed the rmarkdown package  

`install.package("rmarkdown")`

## Open new Markdown file

\begin{center}
\includegraphics[scale=0.45]{IMG/NewRMD}
\end{center}


## Give your file a title

\begin{center}
\includegraphics[scale=0.45]{IMG/NewRMDSet}
\end{center}


## Choose the output format

* You can shoose between HTML, PDF and Docx(Word)  
   + **HTML** is recommeneded, since everybody can read it
* You can always change your output format, in the  YAML Header

## Starting Point

\begin{center}
\includegraphics[scale=0.30]{IMG/RMDDefault}
\end{center}

## YAML Header

\definecolor{green-code}{HTML}{2eb82e}

### Example
\begin{center}
\includegraphics[scale=0.45]{IMG/YAMLHeader}
\end{center}
  
### Output types


Type | Format | Option in YAML Header
-----|--------|------------------------
Website | HTML | \textcolor{green-code}{output}: \textcolor{green-code}{html\_document}  
Document | PDF | \textcolor{green-code}{output}: \textcolor{green-code}{pdf\_document}  
Document (Word) | RTF | \textcolor{green-code}{output}: \textcolor{green-code}{word\_document}  
Presentation (beamer) | PDF |\textcolor{green-code}{output}: \textcolor{green-code}{beamer\_presentation}  
Presentation (ioslides) | HTML | \textcolor{green-code}{output}: \textcolor{green-code}{ioslides\_presentation}  

## Write markdown

You can use the Quick Reference or the Cheatsheet  
Both can be found by opening the **Help** dropdown menu which is positioned in the top bar
\begin{center}
\includegraphics[scale=0.45]{IMG/QuickReference}
\includegraphics[scale=0.35]{IMG/Cheatsheet}
\end{center}

## Syntax Examples

Text formatting |
-----|----------------
Text | Text
`` `r "*italic*"` `` or `` `r "_italic_"` `` | *italic*
`` `r "**bold**"` `` or `` `r "__bold__"` ``| **bold**
`` `r "[link](www.sbg.ac.at)"` `` | [link](www.sbg.ac.at)


## Embed your code
**inline code **   
```{r echo=FALSE}
bq="`"
```
The mean speed of cars is (`` `r sprintf("%sr mean(cars$speed)%s", bq, bq)` ``) `r mean(cars$speed)`. 

**code chunk**   
{r cars, echo=TRUE}

```{r cars, echo=TRUE}
summary(cars)
```

## How to get this " ` " symbol

Press Shift + ` (twice or followed by Space)
\begin{center}
\includegraphics[scale=0.45]{IMG/Keyboard}
\end{center}


## Embed your code
* There are a couple of arguments to customize your output
    + e.g. adding echo = FALSE wilt not show the command in the output
    
## Embed your code
* You can also use other languges like Python, SQL and more to embed your code  

### Example
{python}  
```{python, echo=TRUE}
x = 'hello, python world!'
print(x.split(' '))
```


## Rendering output

\begin{center}
\includegraphics[scale=0.45]{IMG/Knitr}
\end{center}
* Press **"Knit"** in Rstudio 
* Press **Ctrl+Shift+K** (Windows)
* Procedure

\begin{center}
\includegraphics[scale=0.45]{IMG/Procedure}
\end{center}

## References

* https://rmarkdown.rstudio.com/lesson-1.html
* https://www.is.uni-freiburg.de/resources/computational-economics/R_Markdown.pdf
* https://www.youtube.com/watch?v=u4ZdvYXjsIo
* https://bookdown.org/yihui/rmarkdown/


## Exercices

To do the following execices, the R Markdown Reference Guide document will help you.  
  
Please download it from:  
  
https://myfiles.sbg.ac.at/index.php/s/7DrLxa2zZiqz5fH

## Exercices  

You are going to present your work on the "cars" dataset. 

* Title: "Analysis of the cars dataset"  
* Subtitle: Your name  
* Set a header: "Part I: Summary statistics"  
* Write a plain text introduction using empahsis and lists according to Reference Guide 
and reproduce the following:  
  
`I will show you how to use the following commands on the speed variabel in the cars dataset`  
`summary()`  
`mean()`  
`sd()`  

## Exercices

* Create a table that looks as following:

Command | Result
--------|-------
`mean()` | `r mean(cars$speed)`
`sd()` | `r round(sd(cars$speed), 2)`

## Exercices

* Use the `summary()` function to display an overview of the main statistic measures
* Automatically write the Minimum in the output of the sentence
    + `The smallest value of speed is [inline code].`
* Add a horizontal line
* Set a header: "Part II: Visualization"
* Use the command `plot()` to gernerate a scatterplot. Do not print the code!
* Add some plain text, divided into two lines:
    + "Figure I:
    + Correlation of speed and distance"
    
    
## Exercices
* Reproduce this output, with `here` beeing the the following link:
    +  [https://rmarkdown.rstudio.com/](https://rmarkdown.rstudio.com/)  
      
You can learn more about R Markdown [here](https://rmarkdown.rstudio.com/).  

* At the then of your file, add a blockquote:  

> "Stay Healthy!"
